Statistical Machine Learning Models on A Bank Marketing Dataset

### 2025-12-09 Slides.Rmd
slides update7

---
title: "Statistical Machine Learning Models on A Bank Marketing Dataset"
author: "Sandra Afrifa, Tigist Kefelew, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    widescreen: true
---

@@

```{r residuals_slide, echo=FALSE, results='asis'}
residuals <- residuals(glm.baseline, type = "deviance")
train_data$residuals <- residuals

@@

```{r knn-prep, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(class)
library(caret)

@@

```{r knn-tuning, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
set.seed(445)

@@

```{r knn-results, echo=FALSE, message=FALSE, warning=FALSE}
library(pROC)

@@

# 3. Boosting Model (High Predictive Benchmark)

```{r boosting-setup, echo=FALSE, message=FALSE, warning=FALSE}
library(gbm)
library(pROC)
library(caret)

@@

```{r boosting-viz, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
plot(
  boost_model$train.error,
  type = "l",
  xlab = "Number of Trees (Iterations)",
  ylab = "Training Loss",
  main = "Boosting: Sequential Error Reduction",
  col = "steelblue",
  lwd = 2
)
```

@@

```{r boosting-tuning, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# Test different numbers of trees
M_values <- seq(50, 1500, by = 50)
test_error <- numeric(length(M_values))

@@

## ROC Curve Comparison
```{r final-roc, echo=FALSE, message=FALSE, warning=FALSE}
roc_all <- rbind(
  data.frame(fpr = 1 - roc_baseline$specificities, tpr = roc_baseline$sensitivities, Model = "Logistic"),
  data.frame(fpr = 1 - roc_ridge$specificities,    tpr = roc_ridge$sensitivities,    Model = "Ridge"),
  data.frame(fpr = 1 - roc_lasso$specificities,    tpr = roc_lasso$sensitivities,    Model = "LASSO"),
  data.frame(fpr = 1 - roc_knn$specificities,      tpr = roc_knn$sensitivities,      Model = "KNN"),
  data.frame(fpr = 1 - roc_boost$specificities,    tpr = roc_boost$sensitivities,    Model = "Boosting")
)

>>>>

---
title: "Statistical Machine Learning Models on A Bank Marketing Dataset"
author: "Sandra Afrifa, Tigist Kefelew, Mussa Hassen, Waddah Alaahtani"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    widescreen: true
---

@@

```{r residuals_slide, echo=FALSE, fig.align="center", results='asis'}
residuals <- residuals(glm.baseline, type = "deviance")
train_data$residuals <- residuals

@@

```{r knn-prep, echo=FALSE, message=FALSE, fig.align="center", warning=FALSE}
library(dplyr)
library(class)
library(caret)

@@

```{r knn-tuning, echo=FALSE, message=FALSE, fig.align="center", warning=FALSE, cache=TRUE}
set.seed(445)

@@

```{r knn-results, echo=FALSE, message=FALSE, fig.align="center",  out.width="75%", warning=FALSE}
library(pROC)

@@

# 3. Boosting Tress Model

```{r boosting-setup, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
library(gbm)
library(pROC)
library(caret)

@@

```{r boosting-viz, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width="80%"}
plot(
  boost_model$train.error,
  type = "l",
  xlab = "Number of Trees (Iterations)",
  ylab = "Training Loss",
  main = "Boosting: Sequential Error Reduction",
  col = "steelblue",
  lwd = 2
)
```

@@

```{r boosting-tuning, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width="75%", cache=TRUE}
# Test different numbers of trees
M_values <- seq(50, 1500, by = 50)
test_error <- numeric(length(M_values))

@@

## ROC Curve Comparison
```{r final-roc, echo=FALSE, message=FALSE, fig.align="center", warning=FALSE}
roc_all <- rbind(
  data.frame(fpr = 1 - roc_baseline$specificities, tpr = roc_baseline$sensitivities, Model = "Logistic"),
  data.frame(fpr = 1 - roc_ridge$specificities,    tpr = roc_ridge$sensitivities,    Model = "Ridge"),
  data.frame(fpr = 1 - roc_lasso$specificities,    tpr = roc_lasso$sensitivities,    Model = "LASSO"),
  data.frame(fpr = 1 - roc_knn$specificities,      tpr = roc_knn$sensitivities,      Model = "KNN"),
  data.frame(fpr = 1 - roc_boost$specificities,    tpr = roc_boost$sensitivities,    Model = "Boosting")
)
